---
title: "Untitled"
author: "Madeline Abbott"
date: "March 1, 2018"
output: html_document
---

Hi i'm Katja and I contributed! Thanks, Katja!


Some packages...
```{r}
library(readr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(dplyr)
library(survival)
library(geosphere)
```

## Read in data:
```{r}
Taxi <- read_csv("Taxi.csv")
```

### A first look at the data:

### Create a histogram of trip durations
```{r}
ggplot(Taxi, aes(trip_duration)) + geom_histogram(fill = "yellow", col = "black", bins = 200) + scale_x_log10()
```

### A map of pick-up locations
```{r}
states <- map_data("state")
ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = Taxi, aes(x = pickup_longitude, y = pickup_latitude), color = "black", size = 2, alpha = 0.5) 
```
### A map of drop-off locations
```{r}
ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = Taxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 
```


### A map of long trips with duration over 12 hours (43200 seconds)
```{r}
longTaxi <- Taxi %>%
  filter(trip_duration > 43200)

ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = longTaxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 
```

### A map of short trips with duration under 30 seconds
```{r}
shortTaxi <- Taxi %>%
  filter(trip_duration < 30)

ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = shortTaxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 
```



## Things to do:

* remove pick-up and drop-off locations that are far away or in the ocean--check for reasonables travel speeds
* remove trips with extra long durations
* remove trips with really short (like 10 second) durations
* add a rush hour variable
* Kaplan-Meier curves--trip duration by rush hour (maybe time of day), trip duration by passenger number, trip duration by pickup location (airport?), trip duration by day of the week


## Other things (later):

* control for rush hour by day
* neighborhood
* what does store and forward flag mean?


## Data Cleaning:

```{r}
# So far this limits longitude to between -90 and -60 degrees, and latitude to between 37 and 43 degrees,
#cleanTaxi <- Taxi %>%
#  filter(-70 > pickup_longitude) %>% filter(pickup_longitude > -90) %>%
#  filter(37 < pickup_latitude) %>% filter(pickup_latitude < 43) %>%
#  filter(-70 > dropoff_longitude) %>% filter(dropoff_longitude > -90) %>%
#  filter(37 < dropoff_latitude) %>% filter(dropoff_latitude < 43)

# calculate average speed (approximate) and restrict to average speed of over 2 mph and under 70 mph.
#getDistance <- function(location_data) {distGeo(c(location_data[1], location_data[2]), c(location_data[3], #location_data[4]))}
#cleanTaxi$distance_traveled <- apply(cleanTaxi[c("pickup_longitude", "pickup_latitude", "dropoff_longitude", #"dropoff_latitude")], 1, getDistance)  # in meters, distance "as the crow flies"
# convert to miles
#cleanTaxi$distance_traveled <- cleanTaxi$distance_traveled*0.000621371
# get average speed
#cleanTaxi$average_speed <- cleanTaxi$distance_traveled / (cleanTaxi$trip_duration/60/60)
# filter out really slow and really fast taxis
#cleanTaxi2 <- cleanTaxi %>%
#  filter(average_speed > 2) %>% filter(average_speed < 70)

# Restricts duration to under 80000 seconds (about 22 hours) and over 10 seconds
#cleanTaxi3 <- cleanTaxi2 %>%
#  filter(trip_duration > 10) %>%
#  filter(trip_duration < 80000)


# save the cleaner version of the data
#write.csv(cleanTaxi3, file = "cleanTaxi.csv")
```


# Load in clean data
```{r}
cleanTaxi <- read.csv("cleanTaxi.csv")
```

** Make a map of the clean data**
```{r}

# make another map
ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = foo, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 

```





## Preliminary Analysis:

** use cleanTaxi data **

### Plot trip duration with a Kaplan-Meier curve
```{r}
KM_taxi <- survfit(Surv(trip_duration) ~ 1, data = cleanTaxi)
plot(KM_taxi, col = 'purple', xlim = c(0, 20000))
```












