---
title: "Survival Project"
author: "Madeline Abbott, Katja McKiernan, Hayley Stutzman"
date: "March 1, 2018"
output: html_document
---


## Things to do:

* remove pick-up and drop-off locations that are far away or in the ocean--check for reasonables travel speeds
* remove trips with extra long durations
* remove trips with really short (like 10 second) durations
* add a rush hour variable
* Kaplan-Meier curves--trip duration by rush hour (maybe time of day), trip duration by passenger number, trip duration by pickup location (airport?), trip duration by day of the week


## Other things (later):

* control for rush hour by day
* neighborhood
* what does store and forward flag mean?


Some packages...
```{r}
library(readr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(dplyr)
library(survival)
library(geosphere)
library(lubridate)
library(RgoogleMaps)
library(rgdal)
library(rgeos)
library(tmap)
#library(osrm)
library(tidyr)
#library(survminer)
library(data.table)
library(stargazer)


# Cox Snell Residuals
CoxSnell = function(cs,status,xlim=NULL,ylim=NULL)
{
kmcs=survfit(Surv(jitter(cs,amount=(max(cs)-min(cs))/1000),status)~1)$surv
plot(log(-log(kmcs))~sort(log(cs)),xlab="log(Cox-Snell)",ylab="log(-log(S(Cox-Snell)))",xlim=xlim,ylim=ylim)
abline(0,1,col='red')
}
```

## Read in data:
```{r}
Taxi <- read_csv("train.csv")

Taxi <- read.csv("Taxi.csv")
dim(Taxi)
```

### A first look at the data:

### Create a histogram of trip durations
```{r}
ggplot(Taxi, aes(trip_duration)) + geom_histogram(fill = "yellow", col = "black", bins = 200) + scale_x_log10()
```

### A map of pick-up locations
```{r}
states <- map_data("state")
ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = Taxi, aes(x = pickup_longitude, y = pickup_latitude), color = "black", size = 2, alpha = 0.5) 
```
### A map of drop-off locations
```{r}
ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = Taxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 
```


### A map of long trips with duration over 12 hours (43200 seconds)
```{r}
longTaxi <- Taxi %>%
  filter(trip_duration > 43200)

ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = longTaxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 
```

### A map of short trips with duration under 30 seconds
```{r}
shortTaxi <- Taxi %>%
  filter(trip_duration < 30)

ggplot(data = states) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = shortTaxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.5) 
```



## Data Cleaning:

```{r}
# So far this limits longitude to between -90 and -60 degrees, and latitude to between 37 and 43 degrees,
#cleanTaxi <- Taxi %>%
  #filter(-70 > pickup_longitude) %>% filter(pickup_longitude > -90) %>%
  #filter(37 < pickup_latitude) %>% filter(pickup_latitude < 43) %>%
  #filter(-70 > dropoff_longitude) %>% filter(dropoff_longitude > -90) %>%
  #filter(37 < dropoff_latitude) %>% filter(dropoff_latitude < 43)

# calculate average speed (approximate) and restrict to average speed of over 2 mph and under 70 mph.
#getDistance <- function(location_data) {distGeo(c(location_data[1], location_data[2]), c(location_data[3], location_data[4]))}
#cleanTaxi$distance_traveled <- apply(cleanTaxi[c("pickup_longitude", "pickup_latitude", "dropoff_longitude", "dropoff_latitude")], 1, getDistance)                                         # in meters, distance "as the crow flies"
# convert to miles
#cleanTaxi$distance_traveled <- cleanTaxi$distance_traveled*0.000621371
# get average speed
#cleanTaxi$average_speed <- cleanTaxi$distance_traveled / (cleanTaxi$trip_duration/60/60)
# filter out really slow and really fast taxis
#cleanTaxi2 <- cleanTaxi %>%
  #filter(average_speed > 2) %>% filter(average_speed < 70)

cleanTaxi <- read_csv("cleanTaxi.csv")
dim(cleanTaxi)

# Restricts duration to under 7200 seconds (about 2 hours) and over 10 seconds
#cleanTaxi3 <- cleanTaxi %>%
  #filter(trip_duration > 10) %>%
  #filter(trip_duration < 7200)


# save the cleaner version of the data
#write.csv(cleanTaxi3, file = "cleanTaxi.csv")
```


# Load in clean data

**also remove trips with 0 passengers**
```{r}
cleanTaxi <- read.csv("cleanTaxi.csv") %>%
  filter(passenger_count > 0)
```



** Make a map of the clean data**
```{r}
# make another map
states <- map_data("state")
new_york_areas <- subset(states, region %in% c("new york", "new jersey", "connecticut", "pennsylvania"))
new_york_only <- subset(states, region %in% c("new york"))
# drop off
ggplot(data = new_york_only) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = cleanTaxi, aes(x = dropoff_longitude, y = dropoff_latitude), color = "black", size = 2, alpha = 0.3) + xlab("Longitude") + ylab("Latitude") + ggtitle("Drop-off Locations") + coord_map(xlim = c(-75, -72.5),ylim = c(40, 42))
# pick up
ggplot(data = new_york_only) + geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + coord_fixed(1.3) + guides(fill=FALSE) + geom_point(data = cleanTaxi, aes(x = pickup_longitude, y = pickup_latitude), color = "black", size = 2, alpha = 0.3) + xlab("Longitude") + ylab("Latitude") + ggtitle("Pick-up Locations") + coord_map(xlim = c(-75, -72.5),ylim = c(40, 42))

```





# Preliminary Analysis:

** use cleanTaxi data **

## Plot trip duration with a Kaplan-Meier curve
```{r}
KM_taxi <- survfit(Surv(trip_duration) ~ 1, data = cleanTaxi)
plot(KM_taxi, col = 'purple', xlim = c(0, 7200), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration")
```


## Survival curve with weibull distribution
```{r}
Weib_taxi <- survreg(Surv(trip_duration) ~ 1, dist = "weibull", data = cleanTaxi)

curve(1 - pweibull(x, shape = 1/0.7045081, scale = exp(6.83227)), from = 0, to = 23500)

haz_taxi <- function(x) {dweibull(x, shape = 1/0.7045081, scale = exp(6.83227)) / (1-pweibull(x, shape = 1/0.7045081, scale = exp(6.83227)))}

plot(haz_taxi, from = 0, to = 23500)
```



### Find mean trip duration
```{r}
AUCKM = function(survobj,duration)
{
base=c(0,summary(survobj)$time,max(duration))
heights=c(1,summary(survobj)$surv)
new=c()
for(i in 1:length(heights)) { new=c(new,(base[i+1]-base[i])*heights[i]) }
c(sum(new))
}

AUCKM(KM_taxi, cleanTaxi$trip_duration)
```


## Trip duration by passenger number
```{r}
KM_taxi <- survfit(Surv(trip_duration) ~ as.factor(passenger_count), data = cleanTaxi)
plot(KM_taxi, col = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'), xlim = c(0, 23500), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Passenger Count")
legend("topright", inset = 0.05, title = "Number of passengers", legend = c("1", "2", "3", "4", "5", "6"), fill = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'))

# zoom in
plot(KM_taxi, col = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'), xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Passenger Count")
legend("topright", inset = 0.05, title = "Number of passengers", legend = c("1", "2", "3", "4", "5", "6"), fill = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'))

library(stargazer)
stargazer(KM_taxi, align = T)
```

### Log rank test
```{r}
survdiff(Surv(trip_duration) ~ as.factor(passenger_count), data = cleanTaxi)
1 - pchisq((314.117 + 707.589 + 110.091 + 152.807 + 5.734 + 0.522), df = 5)
```

### AFT model
```{r}
m <- survreg(Surv(trip_duration) ~ as.factor(passenger_count), dist = "weibull", data = TaxiRoutes)
summary(m)

surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181))}
integrate(surv, 0, Inf)


# one passenger mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181))}
integrate(surv, 0, Inf)
# two passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181 + 0.0628))}
integrate(surv, 0, Inf)
# three passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181 + 0.0476))}
integrate(surv, 0, Inf)
# four passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181 + 0.0751))}
integrate(surv, 0, Inf)
# five passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181 + 0.0203))}
integrate(surv, 0, Inf)
# six passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.704, scale = exp(6.8181 + 0.0116))}
integrate(surv, 0, Inf)
```

Control for distance and number of steps
```{r}
m2 <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + total_distance + number_of_steps, dist = "weibull", data = TaxiRoutes)
summary(m2)
```

Control for distance
```{r}
m3 <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + total_distance, dist = "weibull", data = TaxiRoutes)
summary(m2)

# one passenger mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.481, scale = exp(6.27 + 9.65e-05 * mean(TaxiRoutes$total_distance)))}
integrate(surv, 0, Inf)
# two passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.481, scale = exp(6.27 + 1.45e-02 + 9.65e-05 * mean(TaxiRoutes$total_distance)))}
integrate(surv, 0, Inf)
# three passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.481, scale = exp(6.27 + 2.69e-02 + 9.65e-05 * mean(TaxiRoutes$total_distance)))}
integrate(surv, 0, Inf)
# four passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.481, scale = exp(6.27 + 3.39e-02 + 9.65e-05 * mean(TaxiRoutes$total_distance)))}
integrate(surv, 0, Inf)
# five passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.481, scale = exp(6.27 + 9.80e-04 + 9.65e-05 * mean(TaxiRoutes$total_distance)))}
integrate(surv, 0, Inf)
# six passengers mean trip time
surv <- function(x) {1-pweibull(x, shape = 1/0.481, scale = exp(6.27 + 0.0116 + 9.65e-05 * mean(TaxiRoutes$total_distance)))}
integrate(surv, 0, Inf)
```

## Trip duration by month
```{r}
TaxiRoutes$month <- unlist(lapply(TaxiRoutes$pickup_datetime, lubridate::month))

library(survival)
KM_month <- survfit(Surv(trip_duration) ~ as.factor(month), data = TaxiRoutes)
plot(KM_month, col = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'), xlim = c(0, 5000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Month")
legend("topright", inset = 0.05, title = "Month", legend = c("1", "2", "3", "4", "5", "6"), fill = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'))
```

### AFT model of duration by month
```{r}
m <- survreg(Surv(trip_duration) ~ as.factor(month), dist = "weibull", data = TaxiRoutes)
summary(m)

# jan
surv <- function(x) {1-pweibull(x, shape = 1/0.702, scale = exp(6.77320))}
integrate(surv, 0, Inf)
# feb
surv <- function(x) {1-pweibull(x, shape = 1/0.702, scale = exp(6.77320 + -0.00355))}
integrate(surv, 0, Inf)
# mar
surv <- function(x) {1-pweibull(x, shape = 1/0.702, scale = exp(6.77320 + 0.03346))}
integrate(surv, 0, Inf)
# apr
surv <- function(x) {1-pweibull(x, shape = 1/0.702, scale = exp(6.77320 + 0.07207))}
integrate(surv, 0, Inf)
# may
surv <- function(x) {1-pweibull(x, shape = 1/0.702, scale = exp(6.77320 + 0.12043))}
integrate(surv, 0, Inf)
# jun
surv <- function(x) {1-pweibull(x, shape = 1/0.702, scale = exp(6.77320 + 0.12584))}
integrate(surv, 0, Inf)

# controlling for distance
m2 <- survreg(Surv(trip_duration) ~ as.factor(month) + total_distance, dist = "weibull", data = TaxiRoutes)
summary(m2)
```

### Month Multivariate Models

#### Parametric models and Comparisons
```{r}
#Exponential
#month_exp = survreg(Surv(trip_duration) ~ as.factor(month) + total_distance, dist = "exponential", data = TaxiRoutes)
#summary(month_exp)

# Weibull:
month_weib = survreg(Surv(trip_duration) ~ as.factor(month), dist = "weibull", data=TaxiRoutes)
summary(month_weib)

monthDist_weib = survreg(Surv(trip_duration) ~ as.factor(month) + total_distance, dist = "weibull", data = TaxiRoutes)
summary(monthDist_weib)

monthWea_weib = survreg(Surv(trip_duration) ~ as.factor(month) + weather, dist = "weibull", data = TaxiWeather)
summary(monthWea_weib)

#Lognormal:
#month_lnorm = survreg(Surv(trip_duration) ~ as.factor(month) + total_distance, dist = "lognormal", data = TaxiRoutes)
#summary(month_lnorm)

#month_exp$loglik
#month_weib$loglik
#month_lnorm$loglik
```

#### Survival curves by month (weibull)
```{r}
curve(1-pweibull(x, shape=1/0.7, scale=exp(6.77352)), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Month")
#curve(1-pweibull(x, shape=1/0.7, scale=exp(6.77352-0.00413)), col = 'orange', add=TRUE)
#curve(1-pweibull(x, shape=1/0.7, scale=exp(6.77352+0.03262)), col = 'black', add=TRUE)
curve(1-pweibull(x, shape=1/0.7, scale=exp(6.77352+0.07169)), col = 'green', add=TRUE)
#curve(1-pweibull(x, shape=1/0.7, scale=exp(6.77352+0.11858)), col = 'cyan', add=TRUE)
curve(1-pweibull(x, shape=1/0.7, scale=exp(6.77352+0.12541)), col = 'purple', add=TRUE)

legend("topright", inset = 0.05, title = "Month", legend = c("1", "4", "6"), fill = c('red', 'green', 'purple'))
```

#### Survival curves of month and distance
```{r}
curve(1-pweibull(x, shape=1/0.479, scale=exp(6.23e+00 + (9.63e-05*mean(TaxiRoutes$total_distance)))), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Month Controlling for Distance")
#curve(1-pweibull(x, shape=1/0.479, scale=exp(6.23e+00+6.33e-03+ 9.63e-05*mean(TaxiRoutes$total_distance))), col = 'orange', add=TRUE)
#curve(1-pweibull(x, shape=1/0.479, scale=exp(6.23e+00+2.01e-02 + 9.63e-05*mean(TaxiRoutes$total_distance))), col = 'black', add=TRUE)
curve(1-pweibull(x, shape=1/0.479, scale=exp(6.23e+00+6.61e-02+ 9.63e-05*mean(TaxiRoutes$total_distance))), col = 'green', add=TRUE)
#curve(1-pweibull(x, shape=1/0.479, scale=exp(6.23e+00+8.76e-02+ 9.63e-05*mean(TaxiRoutes$total_distance))), col = 'cyan', add=TRUE)
curve(1-pweibull(x, shape=1/0.479, scale=exp(6.23e+00+9.94e-02+ 9.63e-05*mean(TaxiRoutes$total_distance))), col = 'purple', add=TRUE)

legend("topright", inset = 0.05, title = "Month", legend = c("1", "4", "6"), fill = c('red', 'green', 'purple'))
```

#### survival curves of month and weather = fair
```{r}
curve(1-pweibull(x, shape=1/0.7, scale=exp(6.78315)), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Month Controlling for Weather = fair")
#curve(1-pweibull(x, shape=1/0.7, scale=exp(6.78315-0.00156)), col = 'orange', add=TRUE)
#curve(1-pweibull(x, shape=1/0.7, scale=exp(6.78315+0.02822)), col = 'black', add=TRUE)
curve(1-pweibull(x, shape=1/0.7, scale=exp(6.78315+0.06753)), col = 'green', add=TRUE)
#curve(1-pweibull(x, shape=1/0.7, scale=exp(6.78315+0.11016)), col = 'cyan', add=TRUE)
curve(1-pweibull(x, shape=1/0.7, scale=exp(6.78315+0.11675)), col = 'purple', add=TRUE)

legend("topright", inset = 0.05, title = "Month", legend = c("1", "4", "6"), fill = c('red', 'green', 'purple'))
```



```{r}
#LRT of Exponential and Weibull
2*(-10399207--10873727)  #t-test
1-pchisq(949040, df = 1)  #p-value

#AIC test for Weibull and Lognormal
2*7 - 2*-10873727 #exp
2*8 - 2*-10399207 #weibull
2*8 - 2*-10417985 #lognormal

#Weibull appears to be the best model!!
```



#### Cox PH models and Testing adequacy
```{r}
#cph models
cph_month = coxph(Surv(trip_duration) ~ as.factor(month), data=TaxiRoutes)
summary(cph_month)
cox.zph(cph_month)

cph_month_dist = coxph(Surv(trip_duration) ~ as.factor(month) + total_distance, data=TaxiRoutes)
summary(cph_month_dist)
cox.zph(cph_month_dist)

cph_month_weather = coxph(Surv(trip_duration) ~ as.factor(month) + weather, data=TaxiWeather)
summary(cph_month_weather)
cox.zph(cph_month_weather)


#cloglog appears to be too close hard to tell?
```

```{r}
plot(KM_month, fun='cloglog', xlim = c(10,50), ylim = c(-10,-6), col=c("red", "orange", "green", "blue", "purple", "black"))
```



## Creating a Rush Hour variable 
```{r}
# Sample data set for adding rush hour
tripIDTest = c(1,2,3,4,5,6,7,8,9,10)
TimeTest = c(16,16, 8,9.5,6,3,1,12, 8.75, 10)
TestTaxi = data.frame(tripIDTest,TimeTest)
TestTaxi
# Function for adding a rush hour variable-this will work once we figure out a good way to extract the time of day from the dataset
CreateRush <- function(data1,timeCat){
  rush = c( 1:nrow(data1))
  for(i in 1:nrow(data1)){
    if(hour(timeCat[i]) <= 8 && hour(timeCat[i]) >= 6 ){
      rush[i] = 1
    }
    else if(hour(timeCat[i]) >= 15 && hour(timeCat[i]) <= 17){
      rush[i] = 2
    }
    else{
      rush[i] = 0
    }
  }
  return(newdata1 = (cbind(data1, rush)))
}

```


```{r}
# Extracting time only from pickup info
cleanTaxi$pickup_datetime = ymd_hms(cleanTaxi$pickup_datetime)
#Adding rush hour variable to cleanTaxi
cleanTaxi = CreateRush(cleanTaxi, cleanTaxi$pickup_datetime)


# Extracting time only from pickup info
TaxiWeather$pickup_datetime = ymd_hms(TaxiWeather$pickup_datetime)
#Adding rush hour variable to cleanTaxi
TaxiWeather = CreateRush(TaxiWeather, TaxiWeather$pickup_datetime)
```


```{r}
KMRush = survfit(Surv(cleanTaxi$trip_duration)~cleanTaxi$rush, conf.type = "plain")

plot(KMRush, col = c("black", "red", "green"), xlim= c(0,5000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Rush Hour")
legend("topright", inset = 0.05, title = "Rush Hour Indicator", legend = c("0", "1", "2"), fill = c('black', 'red', 'green'))

survdiff(Surv(cleanTaxi$trip_duration)~cleanTaxi$rush)

coxRushDist = coxph(Surv(TaxiWeather$trip_duration)~TaxiWeather$rush*TaxiWeather$total_distance)
summary(coxRushDist)

cox.zph(coxRushDist)


KMRushDist = survfit(Surv(TaxiWeather$trip_duration)~TaxiWeather$rush+TaxiWeather$total_distance, conf.type = "plain")

plot(KMRushDist, col = c("black", "red", "green"), xlim= c(0,5000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Rush Hour and Distance")
legend("topright", inset = 0.05, title = "Rush Hour Indicator", legend = c("0", "1", "2"), fill = c('black', 'red', 'green'))
```



# Weather

Join Taxi data with Weather data
```{r}
WeatherData <- read.csv("WeatherTaxi.csv")
WeatherData$date <- format(as.Date(WeatherData$date, format = "%d-%m-%Y"), "%Y-%m-%d")
cleanTaxi$date <- format(as.Date(cleanTaxi$pickup_datetime, format = "%Y-%m-%d  %H:%M"), "%Y-%m-%d")
TaxiWeather <- merge(cleanTaxi, WeatherData, by = "date")

#precipitation
TaxiWeather$precipitation <- as.numeric(as.character(TaxiWeather$precipitation))
TaxiWeather$precipitation[is.na(TaxiWeather$precipitation)] <- 0.01
#snow.fall
TaxiWeather$snow.fall <- as.numeric(as.character(TaxiWeather$snow.fall))
TaxiWeather$snow.fall[is.na(TaxiWeather$snow.fall)] <- 0.01
#snow.depth
TaxiWeather$snow.depth <- as.numeric(as.character(TaxiWeather$snow.depth))
TaxiWeather$snow.depth[is.na(TaxiWeather$snow.depth)] <- 0.01
```


### Trip duration by precipitation

Try modeling with exponential, weibull, and log normal distributions.  Use Cox Snell residuals to compare fits.
```{r}
# Exponential
mod_precip_exp <- survreg(Surv(trip_duration) ~ precipitation, data = TaxiWeather, dist = "exponential")
summary(mod_precip_exp)

# Plot only a sample of 1000 CS residuals
cs = -log( 1 - pexp(TaxiWeather$trip_duration, 1/exp(6.7331 + -0.0394*TaxiWeather$precipitation)))
# smaller sample of cs residuals
cs_sample <- sample(cs, 1000)
status <- rep(1, length(cs_sample))
CoxSnell(cs_sample, status)

# Weibull
mod_precip_weib <- survreg(Surv(trip_duration) ~ precipitation, data = TaxiWeather, dist = "weibull")
summary(mod_precip_weib)
# Plot only a sample of 1000 CS residuals
cs = -log( 1 - pweibull(TaxiWeather$trip_duration, shape = 1/0.704, scale = exp(6.8356 + -0.0392*TaxiWeather$precipitation)))
# smaller sample of cs residuals
cs_sample <- sample(cs, 1000)
status <- rep(1, length(cs_sample))
CoxSnell(cs_sample, status)

# Log normal
mod_precip_lnorm <- survreg(Surv(trip_duration) ~ precipitation, data = TaxiWeather, dist = "lognormal")
summary(mod_precip_lnorm)
# Plot only a sample of 1000 CS residuals
cs = -log( 1 - plnorm(TaxiWeather$trip_duration, mean = 6.4724 + -0.0402*TaxiWeather$precipitation, sd = 0.741))
# smaller sample of cs residuals
cs_sample <- sample(cs, 1000)
status <- rep(1, length(cs_sample))
CoxSnell(cs_sample, status)

# Log normal look the best!


# Trip duration by precipitation (controlling for distance)
summary(survreg(Surv(trip_duration) ~ precipitation + distance_traveled, data = TaxiWeather2), dist = "lognormal")


```

```{r}
#cph_precipitation = coxph(Surv(trip_duration) ~ precipitation, data=TaxiWeather)
#summary(cph_precipitation)

#cox.zph(cph_precipitation)

#Residuals for model had significant p-value so we would not want to use a cox ph model for - stick with the lognormal.
```


### Q: How does the type of weather affect trip duration?

Make new categorical variable: fair weather, rain, snow
```{r}
TaxiWeather <- read.csv("TaxiWeatherDISTANCE.csv")
TaxiWeather <- TaxiWeather %>%
    mutate(weather = ifelse(snow.fall > 0, "snow", ifelse(precipitation > 0, "rain", "fair"))) 
```

Then make models (with and without controlling for distance)
```{r}
weather = survreg(Surv(trip_duration) ~ weather, dist = "lognormal", data = TaxiWeather)
summary(weather)

weather1 = survreg(Surv(trip_duration) ~ weather + total_distance, dist = "lognormal", data = TaxiWeather)
summary(weather1)

weather2 = survreg(Surv(trip_duration) ~ weather * precipitation + total_distance, dist = "lognormal", data = TaxiWeather)
summary(weather2)

weather_dist_passenger = survreg(Surv(trip_duration) ~ weather+distance_traveled+factor(passenger_count), dist = "lognormal", data = TaxiWeather)
summary(weather_dist_passenger)
```

##### Survival curves by weather
```{r}
curve(1-plnorm(x, mean=6.47528, sd=0.741), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Weather Type")

curve(1-plnorm(x, mean=6.47528+0.00368, sd=0.741), col = 'black', add=T)

curve(1-plnorm(x, mean=6.47528-0.07395, sd=0.741), col = 'cyan', add=T)

legend("topright", inset = 0.05, title = "Month", legend = c("fair", "rain", "snow"), fill = c('red', 'black', 'cyan'))
```

##### Survival curves by weather, controllilng for distance
```{r}
curve(1-plnorm(x, mean=6.03+ 9.39e-05 * mean(TaxiWeather$total_distance), sd=0.548), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Weather Type controlling for Distance")

curve(1-plnorm(x, mean=6.03+4.92e-03 + 9.39e-05 * mean(TaxiWeather$total_distance), sd=0.548), col = 'black', add=T)

curve(1-plnorm(x, mean=6.03+-5.73e-02+ 9.39e-05 * mean(TaxiWeather$total_distance), sd=0.548), col = 'cyan', add=T)

legend("topright", inset = 0.05, title = "Month", legend = c("fair", "rain", "snow"), fill = c('red', 'black', 'cyan'))
```


# Taxi Routes

* looking at bike lanes, busy streets, snow priority...

Use shortest road routes between start and end.
```{r}
# shorests routes for each taxi ride
Routes1 <- read.csv("fastest_routes_train_part_1.csv")
Routes2 <- read.csv("fastest_routes_train_part_2.csv")

# join routes with taxi data
RoutesAll <- rbind(Routes1, Routes2)
TaxiRoutes <- merge(cleanTaxi, RoutesAll, by = "id")
```

Separating list of taxi route streets into columns
```{r}

### Attempt at Making variable of busy route or area
#TaxiRoutes$street_for_each_step = sapply(TaxiRoutes$street_for_each_step, FUN = toupper)

#TaxiRoutes <- separate(TaxiRoutes, 'street_for_each_step', paste("street_for_each_step", 1:max(TaxiRoutes$number_of_steps), sep="_"), sep="\\|", extra="drop")

#write.csv(TaxiRoutes, file = "TaxiRoutes.csv")

TaxiRoutes <- fread('TaxiRoutes.csv', header = T, sep = ',') 
#TaxiRoutes <- read.csv("TaxiRoutes.csv")
miniTaxiRoutes <- TaxiRoutes[sample(nrow(TaxiRoutes), 1000), ]
#write.csv(miniTaxiRoutes, "miniTaxiRoutes.csv")
```

Add information about streets (working here...)
```{r}
#Street info:
# (Note: first need to make street names consistent between StreetInfo dataset and TaxiRoutes dataset)
StreetInfo <- fread("nyc_centerline.csv")
View(StreetInfo)

StreetInfo$street <- sub("\\<AVE\\>", "AVENUE", StreetInfo$street)
StreetInfo$street <- sub("\\<ST\\>", "STREET", StreetInfo$street)
StreetInfo$street <- sub("\\<BLVD\\>", "BOULEVARD", StreetInfo$street)
StreetInfo$street <- sub("\\<RD\\>", "ROAD", StreetInfo$street)
StreetInfo$street <- sub("\\<LN\\>", "LANE", StreetInfo$street)
StreetInfo$street <- sub("\\<CT\\>", "COURT", StreetInfo$street)
StreetInfo$street <- sub("\\<PL\\>", "PLACE", StreetInfo$street)
StreetInfo$street <- sub("\\<PKWY\\>", "PARKWAY", StreetInfo$street)
StreetInfo$street <- sub("\\<DR\\>", "DRIVE", StreetInfo$street)
StreetInfo$street <- sub("\\<BRG\\>", "BRIDGE", StreetInfo$street)
StreetInfo$street <- sub("\\<EXPY\\>", "EXPRESSWAY", StreetInfo$street)
StreetInfo$street <- sub("\\<PLZ\\>", "PLAZA", StreetInfo$street)
StreetInfo$street <- sub("\\<LOOP\\>", "LOOP", StreetInfo$street)
StreetInfo$street <- sub("\\<XING\\>", "CROSSING", StreetInfo$street)
StreetInfo$street <- sub("\\<CIR\\>", "CIRCLE", StreetInfo$street)
StreetInfo$street <- sub("\\<TER\\>", "TERRACE", StreetInfo$street)
StreetInfo$street <- sub("\\<XING\\>", "CROSSING", StreetInfo$street)
StreetInfo$street <- sub("\\<APPR\\>", "APPROACH", StreetInfo$street)
StreetInfo$street <- sub("\\<CRES\\>", "CRESCENT", StreetInfo$street)
StreetInfo$street <- sub("\\<SQ\\>", "SQUARE", StreetInfo$street)
StreetInfo$street <- sub("\\<ALY\\>", "ALLEY", StreetInfo$street)
StreetInfo$street <- sub("\\<WAY\\>", "CAUSEWAY", StreetInfo$street)
StreetInfo$street <- sub("\\<WALK\\>", "WALK", StreetInfo$street)
StreetInfo$street <- sub("\\<PARK\\>", "PARK", StreetInfo$street)
StreetInfo$street <- sub("\\<EN\\>", "EN", StreetInfo$street)
StreetInfo$street <- sub("\\<HL\\>", "HILL", StreetInfo$street)
StreetInfo$street <- sub("\\<TRL\\>", "TRAIL", StreetInfo$street)
StreetInfo$street <- sub("\\<TUNL\\>", "TUNNEL", StreetInfo$street)
StreetInfo$street <- sub("\\<HTS\\>", "HEIGHTS", StreetInfo$street)
StreetInfo$street <- sub("\\<FWY\\>", "FREEWAY", StreetInfo$street)
StreetInfo$street <- sub("\\<PT\\>", "POINT", StreetInfo$street)
StreetInfo$street <- sub("\\<GLN\\>", "GLEN", StreetInfo$street)
StreetInfo$street <- sub("\\<TPKE\\>", "TURNPIKE", StreetInfo$street)
StreetInfo$street <- sub("\\<EXT\\>", "EXTENSION", StreetInfo$street)
StreetInfo$street <- sub("\\<PATH\\>", "PATH", StreetInfo$street)
StreetInfo$street <- sub("\\<GDNS\\>", "GARDENS", StreetInfo$street)
StreetInfo$street <- sub("\\<ROAD\\>", "ROAD", StreetInfo$street)
StreetInfo$street <- sub("\\<HWY\\>", "HIGHWAY", StreetInfo$street)
StreetInfo$street <- sub("\\<LK\\>", "LAKE", StreetInfo$street)
StreetInfo$street <- sub("\\<OVAL\\>", "OVAL", StreetInfo$street)
StreetInfo$street <- sub("\\<ROW\\>", "ROW", StreetInfo$street)
StreetInfo$street <- sub("\\<RAMP\\>", "RAMP", StreetInfo$street)
StreetInfo$street <- sub("\\<SLIP\\>", "SLIP", StreetInfo$street)
StreetInfo$street <- sub("\\<VLG\\>", "VILLAGE", StreetInfo$street)
StreetInfo$street <- sub("\\<CLOS\\>", "CLOSE", StreetInfo$street)
StreetInfo$street <- sub("\\<N\\>", "NORTH", StreetInfo$street)
StreetInfo$street <- sub("\\<CRS\\>E", "COURSE", StreetInfo$street)
StreetInfo$street <- sub("\\<RMP\\>", "RAMP", StreetInfo$street)
StreetInfo$street <- sub("\\<GRN\\>", "GREEN", StreetInfo$street)
StreetInfo$street <- sub("\\<OPAS\\>", "OVERPASS", StreetInfo$street)
StreetInfo$street <- sub("\\<VIA\\>", "VIADUCT", StreetInfo$street)
StreetInfo$street <- sub("\\<RDG\\>", "RIDGE", StreetInfo$street)
StreetInfo$street <- sub("\\<BL\\>", "BOULEVARD", StreetInfo$street)
StreetInfo$street <- sub("\\<EXIT\\>", "EXIT", StreetInfo$street)
StreetInfo$street <- sub("\\<ESPL\\>", "ESPLANADE", StreetInfo$street)
StreetInfo$street <- sub("\\<MNR\\>", "MANOR", StreetInfo$street)
StreetInfo$street <- sub("\\<RDWY\\>", "ROADWAY", StreetInfo$street)
StreetInfo$street <- sub("\\<DY\\>", "DY", StreetInfo$street)
StreetInfo$street <- sub("\\<BDG\\>", "BDG", StreetInfo$street)
StreetInfo$street <- sub("\\<CP\\>", "CAMP", StreetInfo$street)

```

```{r}
#streetInfoRoads <- data.frame(streets = sort(unique(StreetInfo$street))) 
#TaxiRoutesStreetsONLY <- TaxiRoutes[,21:66]
#taxiRoads <- data.frame(streets = sort(unique(as.vector(as.matrix(TaxiRoutesStreetsONLY)))))

#taxiRoads$streets <- sub("TH", "", taxiRoads$streets)
#taxiRoads$streets <- sub("ST", "", taxiRoads$streets)
#taxiRoads$streets <- sub("ND", "", taxiRoads$streets)
#taxiRoads$streets <- sub("RD", "", taxiRoads$streets)
#taxiRoads$streets <- sub(" REET", " STREET", taxiRoads$streets)

#missingRoads <- setdiff(streetInfoRoads$streets,taxiRoads$streets)
#setdiff(missingRoads, streetInfoRoads$streets)
# all road names in taxi routes should have street info
```


### Calculating bike lane %

```{r}
# dataframe for name of road traveled in each step
RoadsNameData <- TaxiRoutes[,21:66]
RoadsNameData <- sapply(RoadsNameData, function(x) sub("TH", "", x))
RoadsNameData <- sapply(RoadsNameData, function(x) sub("ST", "", x))
RoadsNameData <- sapply(RoadsNameData, function(x) sub("ND", "", x))
RoadsNameData <- sapply(RoadsNameData, function(x) sub("RD", "", x))
RoadsNameData <- sapply(RoadsNameData, function(x) sub("REET", "STREET", x))

#taxiRoads$streets <- sub("ST", "", taxiRoads$streets)
#taxiRoads$streets <- sub("ND", "", taxiRoads$streets)
#taxiRoads$streets <- sub("RD", "", taxiRoads$streets)
#taxiRoads$streets <- sub(" REET", " STREET", taxiRoads$streets)


mini <- head(RoadsNameData, 100)
RoadsNameData <- sapply(RoadsNameData, function(x) sub("TH", "", x) )

# dataframe for length of road traveled in each step
RoadLengthData <- TaxiRoutes[,67]
RoadLengthData <- separate(RoadLengthData, 'distance_per_step', paste("distance_per_step", 1:max(TaxiRoutes$number_of_steps), sep="_"), sep="\\|", extra="drop")
```


# NEW STUFF--working here
## Trying again with smaller dataset
```{r}
set.seed(1985)
small_ind <- sample(1:nrow(TaxiRoutes), 100)
smallTaxiRoutes <- TaxiRoutes[small_ind,]

RoadsNameData <- smallTaxiRoutes[,21:66]

for (i in 1:nrow(RoadsNameData)){
  if (startsWith(RoadsNameData[i,], "3")) {
    sapply(RoadsNameData[i,], function(x) sub("[^0-9].* ", " ", x), RoadsNameData)
  } 
}



  if(startsWith(RoadsNameData[i,], "3"), sapply(RoadsNameData[i,], function(x) sub("[^0-9].* ", " ", x)), RoadsNameData)

startsWith("100", "1")

#RoadsNameData <- sapply(RoadsNameData, function(x) sub("ST", "", x))
#RoadsNameData <- sapply(RoadsNameData, function(x) sub("ND", "", x))
#RoadsNameData <- sapply(RoadsNameData, function(x) sub("RD", "", x))
#RoadsNameData <- sapply(RoadsNameData, function(x) sub("REET", "STREET", x))
#RoadsNameData <- sapply(RoadsNameData, function(x) sub("SOU", "SOUTH", x))
#RoadsNameData <- sapply(RoadsNameData, function(x) sub("NOR", "NORTH", x))




getBikeLaneTotal <- function() {
  for (i in 1:number_of_steps) {
    
  }
}
```


```{r}
RoadsNameData = read.csv("RoadsNameData.csv")
StreetInfo = read.csv("nyc_centerline.csv")
mini = head(RoadsNameData, 5)
bikeStreetInfo = subset(StreetInfo, bike_lane > 0)
bikeStreetInfo = bikeStreetInfo[,c(3:4,6,8:13)]
bikeStreetInfo = unique(bikeStreetInfo)


for (i in 1:nrow(mini)){
  bikecount = 0
  BikeLaneTripList = vector(mode = "integer",length = nrow(mini))
  for (j in (2:ncol(mini))){
    #if ((mini[i,j]) = NA){
      for (k in 1:nrow(bikeStreetInfo)){
        word = bikeStreetInfo$bike_lane[k]
        if(identical(word, (mini[i,j]))){
            bikecount = bikecount+1
        }
      }
    #}
  }
  BikeLaneTripList[i] = bikecount
}
BikeLaneTripList
```



### Trip Duration by number of turns
```{r}
#TaxiRoutes$number_of_turns = TaxiRoutes$number_of_steps - 1
library(stringr)
#count number of turns - not including roundabounds and on ramps 
TaxiRoutes$turn_num <- str_count(TaxiRoutes$step_maneuvers, "turn")

```

```{r}
ln_step_number1 = survreg(Surv(trip_duration) ~ number_of_turns, dist = "lognormal", data=TaxiRoutes)
summary(ln_step_number1)

ln_step_number2 = survreg(Surv(trip_duration) ~ number_of_turns + distance_traveled, dist = "lognormal", data=TaxiRoutes)
summary(ln_step_number2)

ln_step_number3 = survreg(Surv(trip_duration) ~ as.factor(passenger_count) + number_of_turns, dist = "lognormal", data=TaxiRoutes)
summary(ln_step_number3)
```


### Multivariate Models with Passenger Count

```{r}
# Trying to see how model passenger count with other variables -> Made Cox PH model of duration by passenger count but didn't seem to follow ph assumption. 

#cph_passenger1 = coxph(Surv(trip_duration) ~ as.factor(passenger_count), data = TaxiRoutes)
#summary(cph_passenger1)

#KM1 = survfit(Surv(trip_duration) ~ as.factor(passenger_count), data=TaxiRoutes)
#plot(KM1, fun='cloglog', mark.time=FALSE, col=c("black", "red", "blue", "green", "orange", "purple"))
#cox.zph(cph_passenger1)



#Trying different parametric models 

# Exponential
#pcount_exp <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + total_distance, data = TaxiRoutes, dist = "exponential")
#summary(pcount_exp)

# Weibull
pcount_weib_dist <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + total_distance, data = TaxiRoutes, dist = "weibull")
summary(pcount_weib_dist)

pcount_weib_steps <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + number_of_steps, data = TaxiRoutes, dist = "weibull")
summary(pcount_weib_steps)

pcount_weib_steps_dur <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + number_of_steps + total_distance, data = TaxiRoutes, dist = "weibull")
summary(pcount_weib_steps_dur)


#Lognormal
# Exponential
#pcount_lnorm <- survreg(Surv(trip_duration) ~ as.factor(passenger_count) + total_distance, data = TaxiRoutes, dist = "lognormal")

#summary(pcount_lnorm)

#pcount_exp$loglik
#pcount_weib$loglik
#pcount_lnorm$loglik
```

```{r}
#LRT for exponential and weibull
2*(-10403739--10874473)
1-pchisq(276644, df=1)
#AIC for all
2*6-2*-10874473 #exponential
2*7-2*-10403739 #weibull
2*7-2*-10419616 #lognormal

#Weibull is the best model
```

#### Passenger and distance
```{r}
curve(1-pweibull(x, shape=1/0.481,scale=exp(6.28e+00-0.1738 +9.65e-05*mean(TaxiRoutes$total_distance))), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Passenger Count Controlling for Distance")
#passenger 2
curve(1-pweibull(x, shape=1/0.481,scale=exp(6.28e+00+5.81e-03+9.65e-05*mean(TaxiRoutes$total_distance))), col = 'orange', add=T)
#passenger 3
curve(1-pweibull(x, shape=1/0.481,scale=exp(6.28e+00+1.82e-02+9.65e-05*mean(TaxiRoutes$total_distance))), col = 'gold', add=T)
#passeenger 4
curve(1-pweibull(x, shape=1/0.481,scale=exp(6.28e+00+2.52e-02+9.65e-05*mean(TaxiRoutes$total_distance))), col = 'green', add=T)
#passenger 5
curve(1-pweibull(x, shape=1/0.481,scale=exp(6.28e+00-7.74e-03+9.65e-05*mean(TaxiRoutes$total_distance))), col = 'cyan', add=T)
#passenger 6
curve(1-pweibull(x, shape=1/0.481,scale=exp(6.28e+00+0.00e+00+9.65e-05*mean(TaxiRoutes$total_distance))), col = 'purple', add=T)

legend("topright", inset = 0.05, title = "Passenger Count", legend = c("1", "2", "3", "4", "5", "6"), fill = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'))
```


#### Passenger and number of steps
```{r}
curve(1-pweibull(x, shape=1/0.526,scale=exp(6.1649-0.1738 +0.0978*mean(TaxiRoutes$number_of_steps))), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Passenger Count Controlling for Number os Steps in Route")
#passenger 2
#curve(1-pweibull(x, shape=1/0.526,scale=exp(6.1649-0.1514+0.0978*mean(TaxiRoutes$number_of_steps))), col = 'orange', add=T)
#passenger 3
#curve(1-pweibull(x, shape=1/0.526,scale=exp(6.1649-0.1408+0.0978*mean(TaxiRoutes$number_of_steps))), col = 'gold', add=T)
#passeenger 4
curve(1-pweibull(x, shape=1/0.526,scale=exp(6.1649-0.1314+0.0978*mean(TaxiRoutes$number_of_steps))), col = 'green', add=T)
#passenger 5
#curve(1-pweibull(x, shape=1/0.526,scale=exp(6.1649-0.1690+0.0978*mean(TaxiRoutes$number_of_steps))), col = 'cyan', add=T)
#passenger 6
curve(1-pweibull(x, shape=1/0.526,scale=exp(6.1649-0.1611+0.0978*mean(TaxiRoutes$number_of_steps))), col = 'purple', add=T)

legend("topright", inset = 0.05, title = "Passenger Count", legend = c("1", "2", "3", "4", "5", "6"), fill = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'))
```

#### Passenger and distance and number of steps
```{r}
curve(1-pweibull(x, shape=1/0.479,scale=exp(6.15 + 3.09e-02*mean(TaxiRoutes$number_of_steps) + 7.13e-05*mean(TaxiRoutes$total_distance))), col = 'red', xlim = c(0, 2000), xlab = "Trip Duration (seconds)", ylab = "Survival Estimate", main = "Taxi Trip Duration by Passenger Count Controlling for Distance")
#passenger 2
curve(1-pweibull(x, shape=1/0.479,scale=exp(6.15 + 1.34e-02 + 3.09e-02*mean(TaxiRoutes$number_of_steps) + 7.13e-05*mean(TaxiRoutes$total_distance))), col = 'orange', add=T)
#passenger 3
curve(1-pweibull(x, shape=1/0.479,scale=exp(6.15 + 2.67e-02 + 3.09e-02*mean(TaxiRoutes$number_of_steps) + 7.13e-05*mean(TaxiRoutes$total_distance))), col = 'gold', add=T)
#passeenger 4
curve(1-pweibull(x, shape=1/0.479,scale=exp(6.15 + 3.27e-02  + 3.09e-02*mean(TaxiRoutes$number_of_steps) + 7.13e-05*mean(TaxiRoutes$total_distance))), col = 'green', add=T)
#passenger 5
curve(1-pweibull(x, shape=1/0.479,scale=exp(6.15 + 1.07e-03  + 3.09e-02*mean(TaxiRoutes$number_of_steps) + 7.13e-05*mean(TaxiRoutes$total_distance))), col = 'cyan', add=T)
#passenger 6
curve(1-pweibull(x, shape=1/0.479,scale=exp(6.15 + 9.95e-03 + 3.09e-02*mean(TaxiRoutes$number_of_steps) + 7.13e-05*mean(TaxiRoutes$total_distance))), col = 'purple', add=T)

legend("topright", inset = 0.05, title = "Passenger Count", legend = c("1", "2", "3", "4", "5", "6"), fill = c('red', 'orange', 'gold', 'green', 'cyan', 'purple'))
```



